---
title: \vspace{3in}Tame your data in tidyverse
author: "Seun Odeyemi"
date: "October 13, 2018"
output: 
  pdf_document:
    df_print: tibble
    toc: true
  html_document:
    toc: true
---

\centering
\raggedright
\newpage


```{r setup, echo=TRUE}
knitr::opts_knit$set(root.dir = "C:/Users/USER/tidyverse/datasets")
```

```{r packrat_init, echo=FALSE}
# packrat::init()
```

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
# devtools::session_info()
```

# Load Packages
```{r loadpackages, message=FALSE}
library(tidyverse)
library(skimr)
```

# Cast column types with readr col_types

```{r}
# (bakers_raw <- read_csv("bakers_raw.csv"))
desserts <- read_csv("desserts.csv",
                      col_types = cols(
                        technical = col_number(),
                        uk_airdate = col_date(format = "%d %B %Y"),
                        result = col_factor(levels = NULL)),
                      na = c("", "NA", "N/A")
                     )

desserts %>% dplyr::slice(1:10)
# parse_number("2nd")

# problems(desserts) # a readr function - stop_for_problems() will throw an error if there are any parsing problems 

```

- when did the last episode air in the UK?
```{r arrange_sort_desc}
desserts %>% 
  arrange(desc(uk_airdate))
```

The last episode in the UK was aired on October 31, 2017. 

# Replace values using dplyr's recode

```{r recode}
# Count rows grouping by nut variable
desserts %>% 
    count(signature_nut, sort = TRUE)
    
# Recode filberts as hazelnuts
desserts <- desserts %>% 
  mutate(signature_nut = recode(signature_nut, "filbert" = "hazelnut"))

# Count rows again 
desserts %>% 
    count(signature_nut, sort = TRUE)
```

## Creating dummy variables with recode

Dummy variables are often used in data analysis to bin a variable into one of two categories to indicate the absence or presence of something. Dummy variables take the value `0` or `1` to stand for, for example, loser or winner. Dummy variables are often factor variables as opposed to numeric.

```{r dummy_variable_with_recode}
# desserts <- desserts %>% 
#   mutate(tech_win = if_else(technical == 1, 1, 0))

desserts <- desserts %>% 
  mutate(tech_win = recode_factor(technical, `1` = 1,
                           .default = 0))
head(desserts)

desserts %>% 
  count(technical == 1,tech_win)
```

```{r reorder_colnames_with_dplyr}
desserts %>% select(noquote(order(colnames(desserts))))

# OR

# desserts[,order(colnames(desserts))]
```

