---
title: \vspace{3in}Tame your data in tidyverse
author: "Seun Odeyemi"
date: "October 13, 2018"
output: 
  pdf_document:
    df_print: tibble
    toc: true
  html_document:
    toc: true
---

\centering
\raggedright
\newpage


```{r setup, echo=TRUE}
knitr::opts_knit$set(root.dir = "C:/Users/USER/tidyverse/datasets")
```

```{r packrat_init, echo=FALSE}
# packrat::init()
```

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
# devtools::session_info()
```

# Load Packages
```{r loadpackages, message=FALSE}
library(tidyverse)
library(skimr)
```

# Cast column types with readr col_types

```{r}
# (bakers_raw <- read_csv("bakers_raw.csv"))
desserts <- read_csv("desserts.csv",
                      col_types = cols(
                        technical = col_number(),
                        uk_airdate = col_date(format = "%d %B %Y"),
                        result = col_factor(levels = NULL)),
                      na = c("", "NA", "N/A")
                     )

desserts %>% dplyr::slice(1:10)
# parse_number("2nd")

# problems(desserts) # a readr function - stop_for_problems() will throw an error if there are any parsing problems 

```

- when did the last episode air in the UK?
```{r arrange_sort_desc}
desserts %>% 
  arrange(desc(uk_airdate))
```

The last episode in the UK was aired on October 31, 2017. 

# Replace values using dplyr's recode

```{r recode}
# Count rows grouping by nut variable
desserts %>% 
    count(signature_nut, sort = TRUE)
    
# Recode filberts as hazelnuts
desserts <- desserts %>% 
  mutate(signature_nut = recode(signature_nut, "filbert" = "hazelnut"))

# Count rows again 
desserts %>% 
    count(signature_nut, sort = TRUE)
```

## Creating dummy variables with recode

Dummy variables are often used in data analysis to bin a variable into one of two categories to indicate the absence or presence of something. Dummy variables take the value `0` or `1` to stand for, for example, loser or winner. Dummy variables are often factor variables as opposed to numeric.

```{r dummy_variable_with_recode}
# desserts <- desserts %>% 
#   mutate(tech_win = if_else(technical == 1, 1, 0))

desserts <- desserts %>% 
  mutate(tech_win = recode_factor(technical, `1` = 1,
                           .default = 0))
head(desserts)

desserts %>% 
  count(technical == 1,tech_win)
```

```{r reorder_colnames_with_dplyr}
desserts %>% select(noquote(order(colnames(desserts))))

# OR

# desserts[,order(colnames(desserts))]
```

# Subset your tbl_df with select in dplyr


```{r}
ratings <- read_csv("02.03_messy_ratings.csv",
                    na = c("", NA, NaN))
head(ratings)

ratings <- ratings %>% 
  mutate(bbc = recode_factor(channel, 
                             `Channel 4` = 0, 
                             .default = 1))

# ratings %>% 
#   mutate(viewer_growth = filter(episodes == 10))

ratings <- ratings %>% select(noquote(order(colnames(ratings))))
# parse_date('17-Aug-10')
```

```{r}
# # Select to look at variables to plot next
# ratings %>% 
#   select(series, channel, bbc, viewer_growth)
 
# # Make a filled bar chart
# ggplot(ratings, aes(x = series, y = viewer_growth, fill = bbc)) +
#   geom_col()

# Move channel to front and drop 7-/28-day episode ratings
# ratings %>% 
#   select(channel, everything(), -ends_with("day"))
```

# Reformat variables with the janitor package

```{r}
messy_ratings <- read_csv("messy_ratings2.csv",
                    na = c("", NA, NaN))

head(messy_ratings)
# Load the janitor package
library(janitor)

# Reformat to lower camel case
ratings_new <- messy_ratings %>%
  clean_names(case = "snake")

head(ratings_new)
```

```{r}
# Adapt code to also rename 7-day viewer data
# viewers_7day <- ratings %>% 
#     select(series, viewers_7day_ = ends_with("7day"))

# Create tidy data with 7- and 28-day viewers
# tidy_ratings_all <- ratings2 %>% 
#     gather(episode, viewers, ends_with("day"), na.rm = TRUE) %>% 
#     separate(episode, into = c("episode", "days")) %>%  
#     mutate(episode = parse_number(episode),
#            days = parse_number(days)) 
# 
# tidy_ratings_all %>% 
# 	# Count viewers by series and days
#     count(series, days, wt = viewers) %>%
# 	# Adapt to spread counted values
#     spread(days, n, sep = "_")

tidy_ratings <- ratings %>%
    gather(episode, viewers, -series, na.rm = TRUE) %>%
    mutate(episode = parse_number(episode)) %>% 
    group_by(episode) %>% 
    filter(series == 1 | episode == max(episode)) %>% 
    ungroup()
```

