---
title: \vspace{3in}Explore data with tidyverse
author: "Seun Odeyemi"
date: "October 9, 2018"
output: 
  pdf_document:
    df_print: tibble
    toc: true
  html_document:
    toc: true
---

\centering
\raggedright
\newpage


```{r setup, echo=TRUE}
knitr::opts_knit$set(root.dir = "C:/Users/USER/tidyverse/datasets")
```

```{r packrat_init echo=FALSE}
packrat::init()
```

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```

# Load Packages
```{r loadpackages, message=FALSE}
library(tidyverse)
library(skimr)
```

# Import Data
```{r import_data}
bakeoff <- read_csv("bakeoff.csv",
                    na = c("", "NA", "UNKNOWN"))

bakeoff %>% filter(is.na(showstopper))

# convert the "series" variable type from integer to factor
bakeoff <- bakeoff %>% mutate(series = as.factor(series))

# write_csv(bakeoff, path = "bakers_raw.csv")
```

# Getting summary statistics

## Arrange and glimpse

- Question: On which data date did the first episode of the show air in the US?

```{r arrange_glimpse}
bakeoff %>% arrange(us_airdate) %>% glimpse()
```

- Answer: 2014-12-28

## Summarizing data with the skimr function skim

```{r filter_and_skim, include=FALSE}
bakeoff %>%
  filter(!is.na(us_season)) %>% 
  skim()
```

```{r skim_and_summary}
bakeoff %>% 
  skim() %>%  # no argument needed here
  summary() # no argument needed here
```

# Counting data

- How many distinct series are available in the data?

```{r dplyr_distinct}
bakeoff %>% distinct(series)
```

There are eight distinct series. 

## Counting data with the dplyr function count

- How can we count the number of distinct bakers for each series?

```{r dplyr_count}
bakeoff %>% count(series)

# Using count produces the same output as the code below

bakeoff %>% 
  group_by(series) %>% 
  summarize(n = n())


```

The above table shows us the number of bakers that appeared in each series. 

## Count rows by two variables

Here, I had to create a new variable called `aired_us` based on the conditional statement that checks if `us_season` is populated. If `us_season` is not populated assign _FALSE_ to `aired_us`; if populated assign _TRUE_.

```{r create_var_by_mutate}
bakeoff <- bakeoff %>% mutate(aired_us = 
                     ifelse(is.na(us_season), FALSE,
                            ifelse(us_season != '', TRUE, NA_complex_)))

```

Now, we can go ahead and answer the question: how many series have aired in the US?

```{r count_two_vars}
bakeoff %>% 
  count(aired_us, series)
```

We can see that series 4, 5, 6, and 7 have aired in the US. 

What if we are interested in checking for the proportion of bakers in each series for the whole show? 

```{r count_mutate}
bakeoff %>% 
  count(aired_us, series) %>% 
  mutate(prop_bakers = n/sum(n)) %>% 
  mutate(prop_bakers = sprintf("%0.3f", prop_bakers))
```

The `count` function also ungroups a dataframe for us. For instance, if we group_by then summarize, we will get a surprisly different result from the one we go with the count function. 

```{r}
bakeoff %>% 
  group_by(aired_us, series) %>% 
  summarize(n = n()) %>% 
  mutate(prop_bakers = n/sum(n)) %>% 
  mutate(prop_bakers = sprintf("%0.3f", prop_bakers))
```

Compare this with

```{r}
bakeoff %>% 
  group_by(aired_us, series) %>% 
  summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(prop_bakers = n/sum(n)) %>% 
  mutate(prop_bakers = sprintf("%0.3f", prop_bakers))
```

They both do the same thing, but using count we have a more parsimonious code. 

## Count to roll up a level

```{r count_to_roll_up}
bakeoff %>% 
  count(aired_us, series) %>% 
  count(aired_us)
```

The column name `nn` is the number of series that aired in the US. US viewers have seen half of the series aired in the UK.  

```{r}
bakeoff %>% 
  count(result) 
```

```{r}
bakeoff %>% count(result == "SB")
```

```{r}
# Count the number of rows by series and episode
bakeoff %>% count(series, episode)
```

```{r}
# Add second count by series
bakeoff %>% 
  count(series, episode) %>%
  count(series)
```

nn is the number of episodes in each series. 

```{r}
# Count the number of rows by series and baker
bakers_by_series <- bakeoff %>% 
  count(series, baker)
  
# Print to view
# bakers_by_series
  
# Count again by series
bakers_by_series %>% count(series)
  
# Count again by baker
bakers_by_series %>% count(baker, sort = TRUE)
```

```{r}
ggplot(bakeoff, aes(x = episode)) + 
    geom_bar() + 
    facet_wrap(~series)
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
